Illuminating a Dark Lens : A Type Ia Supernova \\ Magnified by Galaxy Cluster Abell 2744

### 2015-05-16 snTomas.tex
Updated abstract to be <250 words

\begin{abstract}
SN \tomas\ is a Type Ia Supernova (\SNIa) discovered at
$z=1.3457\pm0.0001$ behind the galaxy cluster Abell 2744 ($z=0.308$). This
SN has a projected separation from the cluster core of
$\sim$40\arcsec, closer to the critical lensing line than any previous
cluster-lensed Type Ia SN.  As such, it provides a valuable
opportunity to confront gravitational lens models of galaxy clusters
with the direct measurement of an absolute magnification at the edge
of the strong-lensing region.  We derive a tightly constrained measure
of the peak apparent magnitude, corrected for light curve shape and
extinction.  In a cosmology-independent analysis, we find that \tomas\
is $0.77\pm0.15$ magnitudes brighter than unlensed Type Ia
SNe at similar redshift, implying a lensing magnification
of $\mu_{\rm obs}=2.03\pm0.29$. We compare this direct
measurement of the magnification against predicted magnifications
from 15 lens models developed by 8 independent groups,
representing a broad range of methodologies and incorporating a
variety of strong- and weak-lensing constraints.  The models are
collectively fairly accurate: 8 of the models deliver median
magnifications that are consistent with the measured $\mu$ to within
the 1$\sigma$ uncertainties.  However, the models exhibit a
subtle systematic bias: all of the models that disagree with the SN by
more than $1\sigma$ are {\it higher} than the measured value
and 6 of these models are $\geq1.5\sigma$ above the measured
magnification.  We evaluate possible causes for this mild
bias, and reject explanations that
invoke misinterpretation of the SN data or errors in the
definition of the lens models. We conclude that -- at least for this
cluster and this line of sight -- the most accurate models are those
that employ stringent quality cuts for strong-lensing
constraints. A more robust evaluation of systematic biases
could be achieved with further cooperative lens model comparisons
using a larger sample of lensed Type Ia SNe as well as cluster lensing
simulations.
\end{abstract}

>>>>

\begin{abstract}
\change{Reduced to $<250$ words.}\\
SN \tomas\ is a Type Ia Supernova (SN) discovered at
$z=1.3457\pm0.0001$ behind the galaxy cluster Abell 2744
($z=0.308$). This SN 
%has a projected separation from the cluster core
%of $\sim$40\arcsec, 
is closer to the critical lensing line than any
previous cluster-lensed Type Ia SN, 
%As such, it provides 
providing a valuable
opportunity to test galaxy cluster lens models
with the direct measurement of an absolute magnification at the edge
of the strong-lensing region.  
%We derive a tightly constrained measure
%of the peak apparent magnitude, corrected for light curve shape and
%extinction.  
In a cosmology-independent analysis, we find that \tomas\ is
$0.77\pm0.15$ magnitudes brighter than unlensed Type Ia SNe at similar
redshift, implying a lensing magnification of 
$\mu_{\rm obs}=2.03\pm0.29$. 
\change{In future lens models of Abell 2744 this could
provide valuable leverage for breaking degeneracies that positional
constraints cannot disentangle.}  We compare the measured $\mu$
against predicted magnifications from 15 lens models, \change{ of
which 13 were generated before the SN magnification was known,
qualifying as a pure ``blind test.''}  The models are collectively
fairly accurate: \change{6 of the blind models and both of the post-SN
models} deliver median magnifications that are consistent with the
measured $\mu$ to within 1$\sigma$.  However, there is a subtle
systematic bias: all of the models that disagree with the SN $\mu$ 
%by more than $1\sigma$ 
are {\it higher} than the measured value and 6 of these
are $\geq1.5\sigma$ above the measured magnification.  We
evaluate possible causes for this mild bias,
%and reject explanations
%that invoke misinterpretation of the SN data or errors in the
%definition of the lens models. We 
and conclude that
% -- at least for this
%cluster and this line of sight -- 
the most accurate models are those that employ stringent quality cuts
for strong-lensing constraints. 
% A more robust evaluation
%\change{Further examination} of systematic biases 
%could be achieved 
%\change{could be pursued with more}
%lens model comparisons using lensed
%Type Ia SNe, as well as cluster lensing simulations.
\end{abstract}
